Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-09 manuscript.Rmd
finished distal

```{r part2bins, warning=FALSE, fig.cap="Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
calculate_agreement_bins <- function(ds) {
  ds <- ds %>% group_by(file) %>% 
    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
    add_count(file, bin) %>% ungroup() %>% filter(n > 200) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
                     select(human, file, code, prop_human, bin, num_bins), 
                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
                            pos - factor(pos, levels = pos_levels))
  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
}

@@

difference_plot <- function(ds) {
  ds <- mutate(ds, diff = prop_human - prop_model)
  ggplot(filter_outliers(ds), aes(x = pos, y = diff, fill = pos, group = file)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -.1, ymax = .1, fill = "gray", color = "gray") + 
  #geom_rect(aes(xmin = "Held", xmax = "Upright",ymin = -.1, ymax = .1), fill = "#e5e5e5", color = "#e5e5e5") + 
  stat_summary(data = filter(ds, file %in% outlier_files), position = position_dodge(.4), aes(shape = file), fill = "#333333", size = .75, stroke = .25) +
  stat_summary(position = position_dodge(.4), shape = 21, size = .75, stroke = .25) +
  scale_fill_manual(values = pal, guide = "none") +
  scale_shape_manual(guide = "none", values = 22:23) + 
  coord_cartesian(ylim = c(-1, 1)) + xlab("") + 
  scale_y_continuous(name = "Prediction Difference (min)", breaks = seq(-1, 1, .2), labels = seq(-10, 10, 2), minor_breaks = waiver()) + 
    theme(axis.text.x = element_text(size = 13, color = "black"))
}
difference_plots <- map(agree_bins, ~ difference_plot(.x))
difference_plots <- map2(difference_plots, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
difference_plots[[1]] + difference_plots[[2]]
```

@@

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-min bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)).

>>>>

```{r part2bins, warning=FALSE, fig.cap="Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.", fig.align="center", fig.width=8, fig.height=4.5, cache=TRUE, cache.extra=tools::md5sum("data/compiled_agreement_position.csv"), cache.extra=tools::md5sum("data/compiled_agreement_split.csv")}
calculate_agreement_bins <- function(ds) {
  ds <- ds %>% group_by(file) %>% 
    mutate(sample = row_number(), bin = floor(sample/300)) %>% 
    add_count(file, bin) %>% ungroup() %>% filter(n > 200) %>% group_by(file) %>% mutate(num_bins = length(unique(bin))) %>% ungroup()
  model <- ds %>% count(file, bin, num_bins, pos) %>% add_count(file, bin,num_bins,  wt = n, name = "total") %>% mutate(prop_model = n/total)
  human <- ds %>% count(file, bin, num_bins,  code) %>% add_count(file, bin, num_bins, wt = n, name = "total") %>% mutate(prop_human = n/total)
  agree <- full_join(select(model, file, pos, prop_model, bin, num_bins), 
                     select(human, file, code, prop_human, bin, num_bins), 
                     by = c("file" = "file", "pos" = "code", "bin" = "bin", "num_bins" = "num_bins"))
  agree <- agree %>% mutate(prop_human = ifelse(is.na(prop_human), 0, prop_human),
                            prop_model = ifelse(is.na(prop_model), 0, prop_model),
                            pos - factor(pos, levels = pos_levels))
  agree <- complete(agree, nesting(file, bin, num_bins), pos, fill = list(prop_human = 0, prop_model = 0))
}

@@

difference_plot <- function(ds) {
  ds <- mutate(ds, diff = prop_human - prop_model)
  ggplot(filter_outliers(ds), aes(x = pos, y = diff, fill = pos, group = file)) +
  annotate("rect", xmin = -Inf, xmax = Inf, ymin = -.1, ymax = .1, fill = "gray", color = "gray") + 
  #geom_rect(aes(xmin = "Held", xmax = "Upright",ymin = -.1, ymax = .1), fill = "#e5e5e5", color = "#e5e5e5") + 
  stat_summary(data = filter(ds, file %in% outlier_files), position = position_dodge(.4), aes(shape = file), fill = "#333333", size = .75, stroke = .25) +
  stat_summary(position = position_dodge(.4), shape = 21, size = .75, stroke = .25) +
  scale_fill_manual(values = pal, guide = "none") +
  scale_shape_manual(guide = "none", values = 22:23) + 
  coord_cartesian(ylim = c(-1, 1)) + xlab("") + 
  scale_y_continuous(name = "Prediction Difference (min)", breaks = seq(-1, 1, .2), labels = seq(-10, 10, 2), minor_breaks = waiver()) + 
    theme(axis.text.x = element_text(size = 13, color = "black"))
}
#Difference variable is scaled as proportion, but plotted against min/max of 10 minutes

difference_plots <- map(agree_bins, ~ difference_plot(.x))
difference_plots <- map2(difference_plots, c("A. Group Models", "B. Individual Models"), ~ .x + ggtitle(.y))
difference_plots[[1]] + difference_plots[[2]]
```

@@

```{r prediction-difference}
group_diff <- agree_bins$group %>% 
  mutate(diff = (prop_human - prop_model)*10,
         less_than_1 = ifelse(diff >= -1 & diff <= 1, 1, 0))
group_diff <- group_diff %>% group_by(pos) %>% mutate(n = n()) %>% summarize(prop = sum(less_than_1), total = mean(n)) %>% mutate(prop = apa_num(prop/total*100)) %>% column_to_rownames("pos")

ind_diff <- agree_bins$split %>% 
  mutate(diff = (prop_human - prop_model)*10,
         less_than_1 = ifelse(diff >= -1 & diff <= 1, 1, 0))
ind_diff <- ind_diff %>% group_by(pos) %>% mutate(n = n()) %>% summarize(prop = sum(less_than_1), total = mean(n)) %>% mutate(prop = apa_num(prop/total*100)) %>% column_to_rownames("pos")
```

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)). Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera. Bins were included only if there was > X minutes of usable data.
<!-- RANGE OF BINSm USABLE DATA -->

Table \@ref(tab:pt2binstable) shows the agreement correlation coefficients for group and individual models, including and excluding the outliers identified in the previous section. Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (*r* = `r group_agree_cor_noout`) and individual models (*r* = `r split_agree_cor_noout`). Within-position correlations were weakest for held and strongest for upright regardless of the model type. Agreement for prone was better for group models, whereas sitting and supine were better predicted by individual models.

To describe the observed amount of prediction error in 10-minute bins, we subtracted the predicted duration (in minutes) for each body position in each bin from the human coded duration in that bin to create a prediction difference score. A score of 0 would indicate no error; positive differences indicate that the model overestimated the amount of time in a position, whereas negative differences indicate underestimation. Figure \@ref(fig:part2bins) plots the mean prediction difference for each session for each body position. The gray shaded area marks ± 1 minute of prediction error. The vast majority of session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation (with the previously-identified outliers as exceptions). For group models, we calculated the percentage of 10-min bins across participants that had errors < 1 minute: `r group_diff["Held","prop"]`% for held, `r group_diff["Supine","prop"]`% for supine, `r group_diff["Prone","prop"]`% for prone, `r group_diff["Sitting","prop"]`% for sitting, and `r group_diff["Upright","prop"]`% for upright. For individual models, the percent of 10-min bins with < 1 minute of error was: `r ind_diff["Held","prop"]`% for held, `r ind_diff["Supine","prop"]`% for supine, `r ind_diff["Prone","prop"]`% for prone, `r ind_diff["Sitting","prop"]`% for sitting, and `r ind_diff["Upright","prop"]`% for upright. Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay separating the period that models were trained and tested.

### 2023-03-09 manuscript.tex
finished distal

Figure \ref{fig:part2overall} and Table \ref{tab:part2overalltable} summarize the overall agreement during the distal comparison period. For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \ref{fig:part2overall}). Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded prevalence was \emph{r} = 0.91. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

Visual inspection of \ref{fig:part2overall} indicated two extreme outliers, marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was so much worse compared with the other sessions. The ``gray square'' outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording. In contrast, the ``gray diamond outlier'' had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as ``upright'' but the models predicted it as ``held''. Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.

What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs.~sitting; upright vs.~held) and their accuracy for other classes was strong. This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change). To better capture the typical level of agreement, we report all correlations in Table \ref{tab:part2overalltable} excluding the two outliers. Overall agreement among the non-outlier sessions was excellent for both group models (\emph{r} = 0.95) and individual models (\emph{r} = 0.96). With the exception of held, within-class agreement was strong for the other four body positions.

@@

\caption{Prediction performance (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
\end{figure}

@@

\hypertarget{distal-comparison-agreement-at-a-finer-timescale}{%
\subsubsection{Distal Comparison Agreement at a Finer Timescale}\label{distal-comparison-agreement-at-a-finer-timescale}}

>>>>

Figure \ref{fig:part2overall} and Table \ref{tab:pt2overalltable} summarize the overall agreement during the distal comparison period. For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \ref{fig:part2overall}). Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was \emph{r} = 0.80, and the correlation between individual model predictions and human-coded prevalence was \emph{r} = 0.91. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

Visual inspection of Figure \ref{fig:part2overall} indicated two extreme outliers, which we marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions. The ``gray square'' outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguous appeared during the initial recording. In contrast, the ``gray diamond outlier'' had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as ``upright'' but the models predicted it as ``held''. Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infant's moved while walking upright.

What is notable about both outliers is that disagreements were restricted to a particular border case (supine vs.~sitting; upright vs.~held); accuracy for other classes remained strong. This suggests that their poor performance came as a result of spending a long time in an ambiguous position, not the result of the entire model failing to generalize to the later time period (or an error in sensor placement, such as if the parent removed the leggings and put them on backwards after a diaper change). To better capture the typical level of agreement, we report all correlations in Table \ref{tab:pt2overalltable} excluding the two outliers. Overall agreement among the non-outlier sessions was excellent for both group models (\emph{r} = 0.95) and individual models (\emph{r} = 0.96). With the exception of held, within-class agreement was strong for the other four body positions.

@@

\caption{Prediction difference (difference in minutes between human-coded and model-predicted body position) for 10-minute bins in the distal comparison. Each point shows the mean and SE for a single participant for each body position, summarizing the prediction difference for each of their 10-minute bins. Points falling within the gray shaded region indicate that average prediction errors were less than 1 minute. Performance is plotted separately for (A) group models and (B) individual models. The three outlier participants are plotted in dark gray, with a different shape marking each individual.}\label{fig:part2bins}
\end{figure}

@@

\hypertarget{short-timescale-agreement-during-the-distal-comparison}{%
\subsubsection{Short-Timescale Agreement during the Distal Comparison}\label{short-timescale-agreement-during-the-distal-comparison}}

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated agreement analysis, correlating human-coded body position prevalence with model-predicted prevalence, after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \ref{fig:exemplar-timeline}). Infants had varying numbers of 10-minute bins depending on much time they were awake and on camera. Bins were included only if there was \textgreater{} X minutes of usable data.

Table \ref{tab:pt2binstable} shows the agreement correlation coefficients for group and individual models, including and excluding the outliers identified in the previous section. Performance at a short timescale was similar to performance overall: Overall agreement after excluding outliers was excellent for both group (\emph{r} = 0.92) and individual models (\emph{r} = 0.94). Within-position correlations were weakest for held and strongest for upright regardless of the model type. Agreement for prone was better for group models, whereas sitting and supine were better predicted by individual models.

To describe the observed amount of prediction error in 10-minute bins, we subtracted the predicted duration (in minutes) for each body position in each bin from the human coded duration in that bin to create a prediction difference score. A score of 0 would indicate no error; positive differences indicate that the model overestimated the amount of time in a position, whereas negative differences indicate underestimation. Figure \ref{fig:part2bins} plots the mean prediction difference for each session for each body position. The gray shaded area marks ± 1 minute of prediction error. The vast majority of session-averaged predictions fall within 1 minute of error without a clear bias towards overestimation or underestimation (with the previously-identified outliers as exceptions). For group models, we calculated the percentage of 10-min bins across participants that had errors \textless{} 1 minute: 94.68\% for held, 80.85\% for supine, 94.68\% for prone, 77.66\% for sitting, and 94.68\% for upright. For individual models, the percent of 10-min bins with \textless{} 1 minute of error was: 92.94\% for held, 88.24\% for supine, 88.24\% for prone, 83.53\% for sitting, and 96.47\% for upright. Taken together, these results suggest that most estimates across models, sessions, and body positions had small amounts of prediction error even after a long delay from when models were trained.
